---
title: "data_cleaning"
author: "Fan Gao (Scott)"
date: "2024-01-17"
output: 
  bookdown::html_document2:
    toc: true
---

This is my first (Table \@ref(tab:kable)); this is my first (Figure \@ref(fig:lineplot)), and this is my second (Figure \@ref(fig:Histogram))



# Importing packages 

```{r importing necessary packages}
library(dplyr)
library(readr)
library(tidyverse)
library(ggplot2)
library(stats)
library(knitr)
library(kableExtra)
```

# Counting the number of datas inside BIDS
```{r Counting number of files}
folder_path <- "bids_dataset/Binocular_rivalry"
files <- list.files(folder_path)
print(length(files))
```

# Reading in 54 participants' data
```{r read-and-compile-participants, message = FALSE, warning = FALSE}
# creating a loop to extra all of the participants' information
participant_data <- c()
for(id in 1:58){
  for(session_id in 1:2){
    file_name <- sprintf("sub-%02d_task-rivalry_run-%02d_events", id, session_id)
    file_path <- file.path(folder_path, paste0(file_name, ".tsv"))
# debugging    
#    cat("Checking file:", file_path, "\n")
    if (file.exists(file_path)){
      participant_session_data <- read_tsv(file_path, show_col_types = FALSE)
# adding participant id and session id      
      participant_session_data$subject <- id
      participant_session_data$session <- session_id
# append all of participants information into a single data
      participant_data <- append(participant_data, list(participant_session_data))
      
    }
    #debugging
    else {
      cat("File not found:", file_path, "\n")
      if (id %in% c(43, 53, 54, 56)) {
        cat("Skipping missing participant:", id, "\n")
      }
      next
  }
}}
final <- bind_rows(participant_data)
```

# Counting the number of participants in my new data to double check
```{r confirming the number of participants}
num_participants <- final %>% 
  distinct(subject) %>% 
  n_distinct()

print(num_participants)
```

# Examine your dataset in RStudio. Answer these questions with comments in your read-in chunk
- Tidy? 
 - Yes, each column is a single variable, and they are "onset", "duration", "trail", "sync_side", "key", "dominant", "subject", "session"

 - Yes, each row is a single observation, for example each row is corrsponding to each participant at each moment, wether he/she pressing the key responding to sychronized and unsychronized heartrate. 
 - Yes, each cell is a single measurment, for example under "dominant", each cell is corrsponding to whether the participant pressed the key under sycrhonized or unsychronized heartrate condition. 
- It is tidy enough, I just need to drop some NAs. 
- The dominant, duration, key, participant, sessions are useful variable for this analysis; onset and trails are not useful variables.

```{r write out the data frame to a csv}
write.csv(final, "bids_dataset/compiled_part.csv",row.names = FALSE)
```

# Using dplyr functions
```{r grouping datas, warning = FALSE}
sub_data <- final %>% 
  mutate(duration = as.numeric(duration)) %>% 
  drop_na(duration) %>% 
  group_by(subject,dominant) %>% 
  summarize(total_duration = sum(duration, na.rm = TRUE))
write.csv(sub_data, "bids_dataset/total_durations.csv",row.names = FALSE)
```


# Average duration
```{r average duration}
average_duration <- final %>%
  filter(!(duration == "n/a" | dominant == "n/a")) %>% 
  group_by(subject, sync_side, dominant) %>%
  summarize(average_duration = mean(as.numeric(duration), na.rm = TRUE))
```

# Total duration for each participant
```{r total duration}
filtered_data <- final %>% 
  filter(!(duration == "n/a" | dominant == "n/a")) %>% 
  mutate(duration = as.numeric(duration)) %>%
  data.frame()
# debugging
# has_nas <- any(is.na(filtered_data$duration))
# has_nas
```

### Conducting analysis

```{r paired t-test}
# Example of the t-test
  
  # Perform a t-test comparing 'duration' between two groups of 'dominant'
  t_test_result <- t.test(average_duration ~ dominant, data = average_duration, paired = TRUE)
  
  # Display the t-test results
  print(t_test_result)

# t = -2.5272(p<0.05)  lme4/glmer

```

```{r}
# Assuming your data frame is named df
library(lme4)

# Fit a linear mixed-effects model
lmer_model <- lmer(average_duration ~ dominant + (1 | subject), data = average_duration)

# Print the summary of the model
summary(lmer_model)
```

# Conducting paired t-test
```{r Conducting paired t-test}
average_duration[average_duration$dominant == 'synchronous',]
average_duration[average_duration$dominant == 'synchronous',]$average_duration
sync <- average_duration[average_duration$dominant == 'synchronous',]$average_duration
async <- average_duration[average_duration$dominant == 'asynchronous',]$average_duration
t.test(sync, async, paired = TRUE)
```

# Conducting one way t-test
```{r Conducting one way t-test}
t.test(sync, async, paired = TRUE, alternative = 'greater')
```
# conducting Generalized Linear Mixed-Effects Models dominant + subject

```{r glmer dominant and subject}
glmer(duration ~ dominant + (1 + dominant|subject), 
      family = Gamma(link = 'log'), 
      data = filtered_data)
fit <- glmer(duration ~ dominant + (1 + dominant|subject), 
             family = Gamma(link = 'log'), 
             data = filtered_data)
summary(fit)
```
# conducting Generalized Linear Mixed-Effects Models dominant + subject
```{r glmer subject}
fit1 <- glmer(duration ~ dominant + (1|subject), 
              family = Gamma(link = 'log'), 
              data = filtered_data)
summary(fit1)
```
# conducting Generalized Linear Mixed-Effects Models key + subject
```{r glmer subject+key}
fit2 <- glmer(duration ~ dominant +  key + (1 + key|subject), 
             family = Gamma(link = 'log'), 
             data = filtered_data)
summary(fit2)
```

# ANOVA 
```{r}
anova(fit, fit1, fit2)

```

# Graphing

### line plot
- In this plot, syn and asyn are on the axis. 
- This is a point plot with line plot connect the points
- I am comparing the dominance(syn + asyn) across 54 participants
- I am expecting that the line should overall going up because participants should overall spend more time on syn stimulus
```{r lineplot, fig.cap = "Average Duration for Syn and Asyn"}
ggplot(average_duration, aes(x = dominant, 
                             y = average_duration,
                             group = subject,
                             color = factor(subject)))+
  geom_line(size = .3, alpha = .5) +
  geom_point(size = 3) +
  stat_summary(aes(group = 1), fun = mean, geom = "line", linetype = "solid", color = "black", size = 1) +
  labs(title = "Average duration for syn and asyn",
       x = "Dominance",
       y = "Average Duration",
       subtitle = "n=54") +
  facet_wrap(~sync_side, scales = "free") +
  theme(panel.background = element_rect(fill = 'mistyrose'),
        panel.grid.major = element_line(linetype = 'dashed', color="grey"),
        panel.border = element_rect(colour = "black", fill = NA, linetype = 1),
        axis.text = element_text(colour = "black", 
                                 face = "italic", 
                                 family = "Times New Roman"),
        strip.background = element_rect(fill = "darkgreen"),
        strip.text = element_text(color = "white", hjust = 0),
        text = element_text(family = "Times New Roman"))

```

### Histogram
- The substracted average duration Syn - Asyn are on x-axis, the y axis represents count
- This is a histogram
- I am comparing the difference across 54 participants
- I am anticipate to see more results lies on >0 because when Syn - Asyn should be positive values
```{r Histogram, fig.cap = "Histogram for substracted difference", echo = FALSE}
average_diff <- average_duration %>%
  group_by(subject, sync_side) %>%
  summarise(diff_duration = mean(average_duration[dominant == "synchronous"]) - mean(average_duration[dominant == "asynchronous"]))

ggplot(average_diff, aes(x = diff_duration)) +
  geom_histogram(binwidth = 0.1, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Difference in Average Duration (Syn - Asyn)",
       x = "Difference in Duration",
       y = "Count",
       subtitle = "n=54") +
  facet_wrap(~sync_side, scales = "free") +
  theme(panel.background = element_rect(fill = 'mistyrose'),
        panel.grid.major = element_line(linetype = 'dashed', color="grey"),
        panel.border = element_rect(colour = "black", fill = NA, linetype = 1),
        axis.text = element_text(colour = "black", 
                                 face = "italic", 
                                 family = "Times New Roman"),
        strip.background = element_rect(fill = "darkgreen"),
        strip.text = element_text(color = "white", hjust = 0),
        text = element_text(family = "Times New Roman"))
```
```{r kable, echo = FALSE, fig.cap = "Generalized Linear Mixed-Effects Models key + subject"}
# Extracting information from the GLMM summary
summary_data <- data.frame(
  Effect = c("Intercept", "Dominant", "Key"),
  Estimate = c(0.906608, 0.019277, 0.032863),
  `SEr` = c(0.046824, 0.007174, 0.027494),
  `t-value` = c(19.362, 2.687, 1.195),
  `p-value` = c("< 2e-16", "0.00721**", 0.23198)
)


# Print the table using kable
caption_text = "<span style='font-family: Times; font-size: 16px; font-style: italic;'>Generalized Linear Mixed-Effects Models key + subject</span>"
fit2_kable <- kable(summary_data, 
      align = c("l", "r", "r", "r", "r"),
      escape = FALSE,
      caption = caption_text,
      row_spec = list(c(0:10), background = rep(c("#f7f7f7", "#ffffff"), length.out = 11))) %>%
  kable_styling(position = "left", full_width = F, font_size = 15) %>% 
  column_spec(1:5, width = "1in") %>%
  add_header_above(c(" ", "Estimate and SE" = 2, "Model fit" = 2))

# Add a footnote inside the table
fit2_kable <- fit2_kable %>%
  footnote(general = "*p < 0.05, **p < 0.01, ***p < 0.001")
fit2_kable
```




```{r}


#ggplot(average_duration, aes(x = sync_side, y = average_duration, fill = dominant)) +
#  geom_boxplot() +
#  labs(title = "Syn vs. Average Duration",
#       x = "Syn",
#       y = "Average Duration",
#       fill = "Sync_side") +
# coord_cartesian(ylim = c(0, 6))



```



### Below are my experimenting graphs...
```{r graph}
# Still working on graphing
#dot_plot <- ggplot(average_duration, aes(x = dominant, y = average_duration)) +
#  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 0.5, binwidth = 0.1, position = "jitter") +
#  labs(x = "dominant", y = "duration") +
#  theme_minimal()
#dot_plot

```
```{r}

#ggplot(average_duration, aes(x = dominant, y = average_duration)) +
#  geom_boxplot() +
#  labs(x = "dominant", y = "duration") +
#  theme_minimal()
```

```{r}

#ggplot(average_duration, aes(x = dominant, y = average_duration)) +
#  geom_violin() +
#  labs(x = "dominant", y = "duration") +
#  theme_minimal()
```

```{r}
#ggplot(average_duration, aes(x = dominant, y = average_duration)) +
#  geom_point(position = position_jitter(width = 0.2)) +
#  labs(x = "dominant", y = "duration") +
#  theme_minimal() +
#  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

